Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-02-01 paper/paperBlinded.tex
style changes

\usepackage{etex}
\usepackage[nobiblatex]{helpers/mypackages}
\usepackage[backend=bibtex,natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{choicePrince,biblio}
\usepackage{helpers/myenvironments}
\usepackage{helpers/mycommands}

@@

@@

\pagenumbering{gobble} 
\usepackage[document]{ragged2e}

@@

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

@@

@@

\begin{abstract}
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals for decision making. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

In full generality, a choice mechanism associates decision situations with action choices. A crucial
part of a choice mechanism is the \emph{subjective representation} of the decision situation,
in particular the manner of forming preferences and beliefs about a possibly uncertain world (see
Figure~\ref{fig:ChoiceMechanism} below). This paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness?
The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. A contribution of this paper is
to provide a case that questions of preference evolution should take
variability in uncertainty representation into account as well. As a point in case, we demonstrate
that if agents have imprecise probabilistic beliefs \citep[e.g.][]{gardsah82,levi74,walley96},
preference representations in terms of regret can be selected for, even though they may deviate
from the true, objective fitness that evolutionary selection operates on.

@@

\section{Rationality, representations \& meta-games}
\label{sec:rati--subj}

@@

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. Adaptationism
aims to explain rational behavior by appeal to evolutionary considerations: DMs have acquired
choice mechanisms that have proved to be adpative with respect to the variable environment
where they have evolved. A choice mechanism can be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to closely relate to the literature on evolution of preferences and to the philosophical debate
about the nature of rational choice in general, we here suggest to think of a choice mechanism
as a map from choice situations to action choices which includes an explicit level of
subjective representation of the situation, that we call \emph{subjective representation
  scheme} (see Figure~\ref{fig:ChoiceMechanism}). Specifically, a subjective representation
scheme is a general way of forming preferences and beliefs about the choice situation, and we
are most interested in the question which subjective representations are better than others
from an evolutionary point of view.

@@

\section{Classic choice rules and choice mechanisms}
\label{sec:basic-notions}

@@

The case study under consideration strips the generality of a meta-game approach down to a perspicuous
comparison between four choice mechanisms, summarized in Table~\ref{tab:CMs}.
% the cross-product of two different preference representations (veridical vs.~regret) and two
%different belief representations (precise vs.~imprecise). 
\begin{table}[t]
  \centering
  \begin{tabular}{cccc}
  \multicolumn{3}{c}{choice mechanism} \\  \cmidrule(r){1-3}
    \multicolumn{2}{c}{subjective representation} & action selection & corresponding classic rule \\  \cmidrule(r){1-2}
    preference type & belief type  \\ \midrule
    objective fitness & precise & maxmin EU &  Laplace rule \\
    objective fitness & imprecise & maxmin EU & maxmin \\ 
    regret & precise & maxmin EU & Laplace rule \\
    regret & imprecise & maxmin EU & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent)
    classic choice rules.}
  \label{tab:CMs}
\end{table}
\iffalse
This exclusive selection of choice mechanisms is motivated by a number of
considerations. Firstly, the set of conceivable choice mechanisms is vast. Most of that
conceivable vastness is utter nonsense, and some selection of plausible candidates is required in
any case. Secondly, a restriction to only a small number of suitable candidates will
help demonstrate more distinctly how a meta game approach may help structure philosophical
inquiry, which is our main goal. Thirdly, our case study adds to the recent literature on the
evolution of preferences by making evident the possibly non-trivial interaction between
preference and belief representations, all else equal. This is our secondary goal. 
Finally,
\fi 
The
four choice mechanisms,  
%that we look at here
although phrased in terms of preference and belief representations, map onto well-known classic
proposals for rational decision making (see Table~\ref{tab:CMs}). In other words, there is also
motivation from historical relevance for our selection. To see this, we first recapitulate the
relevant classics ---to have a handy label, we call them \emph{classic rules}--- and then show
how they relate to our notion of choice mechanism.

\subsection{Classic rules for decision making}
\label{sec:choice-rules}

@@

A strategic-form game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i: \bigtimes_{j \in N} A_j \rightarrow \mathbb{R}$ a payoff function for each player $i \in N$.
Part of this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(a_k, a_j) = \pi_2(a_j, a_k)$ for all $a_j,a_k \in A_1$. An example is given in
Figure~\ref{coordgame1}. The numbers give the payoff for the player choosing a row act against
a co-player choosing a column act. Since the game is symmetric, it suffices to give just these
numbers.

Rational choice from maximization of expected utility, as given in
Definition~\ref{def:rationality}, applies to games as follows. To say which acts are rational,
it requires specifying a precise probabilistic belief of DM. Let $A_{-i} := \bigtimes_{j \neq i} A_j$ be
the set of all \emph{profiles} of acts for all players except $i$. If player $i$ has a belief
$\mu_i \in \Delta(A_{-i})$ about the behavior of his co-players, he chooses rationally only if
his choice $a^*_i$ maximizes expected utility
$\EU(a_i, \mu) := \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. For
the example in Figure~\ref{coordgame1}, an expected utility maximizer would choose act $I$ only if he
believes that his co-player is going to choose act $I$ with a probability of at least
$\nicefrac{2}{3}$.

>>>>

\usepackage{etex}

\RequirePackage{amsmath}            % Formeln
\RequirePackage{amsfonts}           % Fonts for Formulas
\RequirePackage{amssymb}
\RequirePackage{amsthm}
\RequirePackage{dsfont}             % double stroke fonts
\RequirePackage[final]{graphicx}
\RequirePackage{booktabs}
\RequirePackage{enumerate}
\RequirePackage{paralist}
\RequirePackage[all]{xy}            
\RequirePackage{url}

\RequirePackage{txfonts} % for strict implication symbols
\RequirePackage{soul}
\RequirePackage{relsize} % provides command \relsize{+/-x} for relative
                     % font size changes
\RequirePackage[german,english]{babel}
\RequirePackage[utf8]{inputenc}
\RequirePackage[T1]{fontenc} 
\RequirePackage{xypic}
\RequirePackage{multicol}
\RequirePackage{subcaption}
\RequirePackage{tikz}
\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}
\RequirePackage{units}

\RequirePackage{dialogue}

\RequirePackage{setspace}

\RequirePackage[colorinlistoftodos,color=lightgray,bordercolor=blue,textsize=footnotesize]{todonotes}

\RequirePackage{xspace} % for \xspace in definition of acronyms etc.

\RequirePackage[final,            % override "draft" which means "no nothing"
            colorlinks,       % rather than outlining them in boxes
            linkcolor=black,   % override truly awful colour choices
            citecolor=black,   %   (ditto)
            urlcolor=black,    %   (ditto)
            plainpages=false, % to overcome complaints with multiple
            pdfpagelabels,    % multiple page 1-s due to preface
            hypertexnames=false % solves warning, but interferes with
                                % index and \autoref apparently
            ]{hyperref}

\usepackage[backend=bibtex,natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}

\bibliography{choicePrince,biblio}

\newtheoremstyle{Satz}
   {}                      %Space above
   {1em}                  %Space below
   {\normalfont}           %Body font
   {}                      %Indent amount (empty = no indent,
                           %\parindent = para indent)
   {\normalfont}           %Thm head font
   {.}                     %Punctuation after thm head
   {.8em}                  %Space after thm head: " " = normal interword
                           %space; \newline = linebreak
   {\bfseries\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}
                           %Thm head spec (can be left empty, meaning
                           %`normal')

\theoremstyle{Satz}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{fact}{Fact}
\newtheorem{claim}{Claim} 
\newtheorem{remark}[theorem]{Remark}
\newtheorem{exercise}{Exercise} 
\newtheorem{problem}{Problem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

\newtheoremstyle{Bsp}
   {}                      %Space above
   {1em}                  %Space below
   {\itshape}           %Body font
   {}                      %Indent amount (empty = no indent,
                           %\parindent = para indent)
   {\normalfont}           %Thm head font
   {.}                     %Punctuation after thm head
   {.8em}                  %Space after thm head: " " = normal interword
                           %space; \newline = linebreak
   {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}
                           %Thm head spec (can be left empty, meaning
                           %`normal')
\theoremstyle{Bsp}
% \newtheorem{example}[theorem]{Example}

% Math --------------------
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\tuple}[1]{\left \langle #1\right\rangle}
\newcommand{\card}[1]{\left \lvert \, #1 \, \right\rvert}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\setbar}{\ensuremath{\thinspace \mid \thinspace}}
\newcommand{\probbar}{\ensuremath{\mid}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\df}{\rightarrow}
\newcommand{\es}{\emptyset}
\newcommand{\den}[1]{\left [\! \left [ #1 \right ]\! \right]}
% \newcommand{\den}[1]{\left \llbracket #1  \right \rrbracket} % this would use fourier package which makes 'cases' environment bad
\newcommand{\no}{\noindent}
\newcommand{\hin}{"$\Rightarrow$" }
\newcommand{\rueck}{"$\Leftarrow$" }
\newcommand{\exs}{\vspace{.15cm}}
\newcommand{\pow}[1]{\ensuremath{\mathcal{P}(#1)}}	% Powerset
\newcommand{\restr}{{\restriction}}
\newcommand{\implicates}{\ensuremath{\leadsto}}  % arrow for
                                % implicatures in examples
\newcommand{\update}[2]{\ensuremath{#1[#2]}}
\newcommand{\myts}{\ensuremath{\thinspace}}
\newcommand{\mycolon}{\ensuremath{\thinspace \colon \thinspace}}
\newcommand{\mydot}{\ensuremath{\thinspace . \thinspace}}

\makeatletter
\newcommand{\prob}{\@ifstar
  \simpleprob%
  \condprob%
}
\def\simpleprob(#1){\ensuremath{\Pr(#1)}}
\def\condprob(#1|#2){\ensuremath{\Pr(#1 \,|\, #2)}}
\makeatother

% General Text Markup--------------------
\newcommand{\runex}[1]{\begin{center}#1\end{center}}
\newcommand{\mydef}[1]{\textsc{#1}}  		% definitions
\newcommand{\markdef}[1]{\textsc{#1}}  		% definitions; alternative
\newcommand{\myment}[1]{\emph{#1}}  	% first mentions
\newcommand{\myword}[1]{\textbf{\texttt{#1}}}	% refering to the word
\newcommand{\myemph}[1]{\emph{#1}}			% emphasis
\newenvironment{exnonum}{
  \begin{list}{}{
    \setlength{\leftmargin}{2.5em}
    \setlength{\rightmargin}{2.5em}
    %\setlength{\itemindent}{-1.5em}
  }
}{
  \end{list}
}

% Slanted Fractions
\newcommand{\myslantfrac}[2]{\msf{#1}{#2}}
\newcommand{\msf}[2]{\ensuremath{\nicefrac{#1}{#2}}}
\newcommand{\msftext}[2]{\nicefrac{#1}{#2}}

% Symbols for Conditionals
\newcommand{\cond}{\ensuremath{>}}
\newcommand{\bicond}{\ensuremath{\Leftrightarrow}}
\newcommand{\strcondWill}{\ensuremath{\boxRight}}
\newcommand{\strcondMight}{\ensuremath{\DiamondRight}}

% Signaling Games & IBR
\newcommand{\sen}{\ensuremath{S}\xspace}		% Sender variable
\newcommand{\mysen}[1]{\ensuremath{\sen^{#1}}} % Sender of type XYZ
\newcommand{\rec}{\ensuremath{R}\xspace}		% Receiver variable
\newcommand{\myrec}[1]{\ensuremath{\rec_{#1}}} % Receiver of type XYZ
\newcommand{\States}{\ensuremath{T}\xspace}		% Set of States
\newcommand{\state}{\ensuremath{t}\xspace}		% single states
\newcommand{\mystate}[1]{\ensuremath{\state_{\text{#1}}}\xspace} %meaningful states
\newcommand{\Messgs}{\ensuremath{M}\xspace}		% Set of Messages
\newcommand{\messg}{\ensuremath{m}\xspace}		% single messages
\newcommand{\mymessg}[1]{\ensuremath{\messg_{\text{#1}}}\xspace} %meaningful messages
\newcommand{\cost}{\ensuremath\operatorname{C}} % cost function
\newcommand{\Acts}{\ensuremath{A}\xspace}		% Set of R-actions
\newcommand{\act}{\ensuremath{a}\xspace}		% single action
\newcommand{\myact}[1]{\ensuremath{\act_{\text{#1}}}\xspace} %meaningful
\newcommand{\Worlds}{\ensuremath{W}}		% Worlds
\newcommand{\world}{\ensuremath{w}}		% single world
\newcommand{\myworld}[1]{\ensuremath{\world_{\text{#1}}}} %named world
\newcommand{\util}{\ensuremath{\operatorname{U}}}	% Utility function
\newcommand{\Util}{\ensuremath{\operatorname{U}}}	% Utility function
\newcommand{\utils}{\ensuremath{\operatorname{U}}}	% Utility function
\newcommand{\Utils}{\ensuremath{\operatorname{U}}}	% Utility function
\newcommand{\RealUtil}{\ensuremath{\operatorname{V}}}	% material payoffs
\newcommand{\Sstrat}{\ensuremath{\sigma}} % Behav/Probab Sender strategy
\newcommand{\Sstrats}{\ensuremath{\mathcal{S}}}	% Set of S-strategies
\newcommand{\Spure}{\ensuremath{s}} % Pure sender strategy
\newcommand{\Spures}{\ensuremath{\mathsf{S}}} % Set of pure sen strategies
\newcommand{\Smixed}{\ensuremath{\tilde{s}}} % Mixed sender strategy
\newcommand{\Smixeds}{\ensuremath{\Delta(\Messgs^\States)}}
\newcommand{\SpuresW}{\ensuremath{\mathsf{S}}} 
\newcommand{\SpuresS}{\ensuremath{\mathsf{S}^{{\mathrm{S}}}}}
\newcommand{\Rstrat}{\ensuremath{\rho}} % Behav/Probab Receiver strategy
\newcommand{\Rstrats}{\ensuremath{\mathcal{R}}}	% Set of R-Strategies
\newcommand{\Rpure}{\ensuremath{r}} % Pure receiver strategy
\newcommand{\Rpures}{\ensuremath{\mathsf{R}}} % Set of pure rec strategies
\newcommand{\Rmixed}{\ensuremath{\tilde{r}}} % Mixed receiver strategy
\newcommand{\Rmixeds}{\ensuremath{\Delta(\Acts^\Messgs)}} 
\newcommand{\RpuresW}{\ensuremath{\mathsf{R}}} 
\newcommand{\RpuresS}{\ensuremath{\mathsf{R}^{\mathrm{S}}}} 
\newcommand{\PureBR}{\ensuremath{\operatorname{BR}}} % Set of pure best responses
\newcommand{\ProbBR}{\ensuremath{\operatorname{BR_{Prob}}}} % Set of mixed best responses
\newcommand{\bel}{\ensuremath{\pi}}
\newcommand{\Bels}{\ensuremath{\Pi}}
\newcommand{\Sbel}{\ensuremath{\pi_{\sen}}}
\newcommand{\Sbels}{\ensuremath{\Pi_{\sen}}}
\newcommand{\Rbel}{\ensuremath{\pi_{\rec}}}
\newcommand{\Rbels}{\ensuremath{\Pi_{\rec}}}
\newcommand{\EU}{\ensuremath{\operatorname{EU}}} % Expected Utility
\newcommand{\EV}{\ensuremath{\operatorname{EV}}} % Expected Response Utility
\newcommand{\BR}{\ensuremath{\operatorname{BR}}} % Best Response
\newcommand{\QR}{\ensuremath{\operatorname{QR}}} % Quantal Response
\newcommand{\WBR}{\ensuremath{\text{{\relsize{-1}W}BR}}} % Weak Best Response
\newcommand{\SBR}{\ensuremath{\text{{\relsize{-1}W}BR}}} % Strong Best Response
\newcommand{\interpr}{\ensuremath{\delta}} % Interpretation strategy

% OT Stuff
\newcommand{\Gen}{\ensuremath{\operatorname{Gen}}} % Generator
\newcommand{\Eval}{\ensuremath{\operatorname{Eval}}} % Evaluator
\newcommand{\Con}{\ensuremath{\operatorname{Con}}} % Constraints
\newcommand{\metsuc}{\ensuremath{\succ}} % Symbol for BiOT metric
\newcommand{\metsuceq}{\ensuremath{\succeq}} % Symbol for BiOT metric
\newcommand{\Go}[1]{\operatorname{Pool}_{#1}}
\newcommand{\Oo}[1]{\operatorname{Opt}_{#1}}
\newcommand{\Bo}[1]{\operatorname{Blo}_{#1}}
\newcommand{\Gr}[1]{\operatorname{GAM^{\rho}}_{#1}}
\newcommand{\Or}[1]{\operatorname{OPT^{\rho}}_{#1}}
\newcommand{\Br}[1]{\operatorname{BLO^{\rho}}_{#1}}

% Abbreviations/Acronyms:
\newcommand{\acro}[1]{\textsc{#1}\xspace}
\newcommand{\acros}[1]{\textsc{#1}{\relsize{-1}s}\xspace}
\newcommand{\bc}{\acro{bc}} % Biscuit Conditional(s)
\newcommand{\bcs}{\acros{bc}}
\newcommand{\cbc}{\acro{cbc}} % Counterfactual BCs
\newcommand{\cbcs}{\acros{cbc}}
\newcommand{\ibr}{\acro{ibr}} % IBR model
\newcommand{\iqr}{\acro{iqr}} % IQR model
\newcommand{\rsa}{\acro{rsa}} % RSA model
\newcommand{\ot}{\acro{ot}} % BiOT
\newcommand{\biot}{\acro{b{\relsize{-1}i}ot}} % BiOT
\newcommand{\tom}{\acro{t{\relsize{-1}o}m}} % ToM
\newcommand{\fc}{\acro{fc}} % Free Choice
\newcommand{\cp}{\acro{cp}} % Conditional Perfection
\newcommand{\uc}{\acro{uc}} % Unconditional Readings
\newcommand{\pbe}{\acro{pbe}}   % Perfect Bayesian Equil.
\newcommand{\pbes}{\acros{pbe}}
\newcommand{\forind}{\acro{fi}} % forward induction
\newcommand{\tcp}{\acro{tcp}} % truth ceteris paribus
\newcommand{\cmr}{\acro{cmr}} % credible message rationalizability
\newcommand{\cm}{\acro{cm}} % credible message (profile) (Rabin)
\newcommand{\condition}[2]{\acro{#1}{#2}} % conditions 
\newcommand{\br}{\acro{br}} % best response (property)
\newcommand{\wbr}{\acro{{\relsize{-1}w}br}} % weak best response (property)
\newcommand{\sbr}{\acro{{\relsize{-1}s}br}} % strong best response (property)
\newcommand{\curb}{\acro{curb}} % curb sets
\newcommand{\gtp}{\acro{gtp}} % game theoretic pragmatics
\newcommand{\sda}{\acro{sda}} % simplification of disjunctive antecedents
\newcommand{\decprob}{\ensuremath{\mathcal{D}}}
\newcommand{\ques}{\ensuremath{\mathfrak{Q}}}
\newcommand{\vsi}{\acro{vsi}}
\newcommand{\evsi}{\acro{evsi}}
\newcommand{\uv}{\acro{uv}}
\newcommand{\qud}{\acro{qud}}
\newcommand{\NE}{\acro{ne}}
\newcommand{\NEs}{\acros{ne}}
\newcommand{\SNE}{\acro{sne}}
\newcommand{\SNEs}{\acros{sne}}
\newcommand{\SG}{\acro{sg}}
\newcommand{\SGs}{\acros{sg}}
\newcommand{\KO}{\textsc{ko\relsize{-1}bs}} % Kennedy's observation
\newcommand{\EVP}{\acro{evp}}    % extreme-value principle
\newcommand{\illc}{\acro{illc}}

% Evolution
\newcommand{\EGT}{\acro{egt}}
\newcommand{\ESS}{\acro{ess}}
\newcommand{\ESSs}{\acros{ess}}
\newcommand{\NSS}{\acro{nss}}
\newcommand{\NSSs}{\acros{nss}}
\newcommand{\sigsys}{\textsc{SigSys}\xspace} % signaling system
\newcommand{\sigsyss}{\textsc{SigSys{\relsize{-1}s}}\xspace} % signaling system Plural

\newcommand{\fin}{\rule{0mm}{1mm}\hfill{\rule{1.5cm}{0.2pt}}}

%Properly typeset tilde for URLs
\def\urltilde{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}

% Beamer footnote for references:
\newcommand{\beamfn}[1]{
  \vfill
      \begin{footnotesize}
        \leftskip 0.1in
        \parindent -0.1in
       \hspace{-0.3cm}\rule{2cm}{0.01cm}\\ \vspace{-0.15cm}
        #1
      \end{footnotesize}
}

\newcommand{\myvec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\transpose}[1]{\ensuremath{\operatorname{T}(#1)}}
\newcommand{\normalize}[1]{\ensuremath{\operatorname{N}(#1)}}

\newcommand{\dn}[1]{\draftnote{#1}}
\newcommand{\fn}[1]{\footnote{#1}}

\newcommand{\stateunmarked}{\ensuremath{\state}\xspace}
\newcommand{\statemarked}{\ensuremath{\state^*}\xspace}
\newcommand{\messgunmarked}{\ensuremath{\messg}\xspace}
\newcommand{\messgmarked}{\ensuremath{\messg^*}\xspace}
\newcommand{\actunmarked}{\ensuremath{\act}\xspace}
\newcommand{\actmarked}{\ensuremath{\act^*}\xspace}

\newcommand{\sunmarked}{\ensuremath{\state}\xspace}
\newcommand{\smarked}{\ensuremath{\state^*}\xspace}
\newcommand{\munmarked}{\ensuremath{\messg}\xspace}
\newcommand{\mmarked}{\ensuremath{\messg^*}\xspace}
\newcommand{\aunmarked}{\ensuremath{\act}\xspace}
\newcommand{\amarked}{\ensuremath{\act^*}\xspace}

\newcommand{\ssome}{\mystate{\exists\neg\forall}}
\newcommand{\sall}{\mystate{\forall}}
\newcommand{\msome}{\mymessg{some}}
\newcommand{\mall}{\mymessg{all}}
\newcommand{\asome}{\myact{\exists\neg\forall}}
\newcommand{\aall}{\myact{\forall}}

% for repeating examples with gb4e

\newcounter{myexememory}
\newenvironment{exer}[1]
{
\setcounter{myexememory}{\value{exx}}
\setcounter{exx}{\getrefnumber{#1}}
\addtocounter{exx}{-1}  
\begin{exe}
}
{
\end{exe}
\setcounter{exx}{\value{myexememory}}
}

\newenvironment{nakedlist}{
  \begin{list}{\quad}{}
}
{
  \end{list}
}

\DefineNamedColor{named}{mycol}{cmyk}{0.6,0.6,0,0}
\DefineNamedColor{named}{mygray}{cmyk}{0.05,0.05,0.05,0.05}
\DefineNamedColor{named}{mygraylight}{cmyk}{0.017,0.017,0.017,0.017}
\DefineNamedColor{named}{mycol2}{cmyk}{0.8,0,0.8,0.2}

\newcommand{\mymark}[1]{{\color{mycol}{#1}}}

\DeclareMathOperator{\expo}{exp}

\newcommand{\mygray}[1]{{\textcolor{gray}{#1}}}
\newcommand{\mycol}[1]{{\textcolor{mycol}{#1}}}
\newcommand{\mycolh}[1]{{\textcolor{mycol2}{#1}}}

% \usepackage{calc}
\newsavebox\CBox
\newcommand\msout[2][0.5pt]{%
  \ifmmode\sbox\CBox{$#2$}\else\sbox\CBox{#2}\fi%
  \makebox[0pt][l]{\usebox\CBox}%  
  \rule[0.5\ht\CBox-#1/2]{\wd\CBox}{#1}}

\newcommand{\greensquare}{\raisebox{1.5pt}{\textcolor{Green}{\Large{\ensuremath{\blacksquare}}}}}
\newcommand{\bluecircle}{\textcolor{blue}{\Huge{\ensuremath{\bullet}}}}
\newcommand{\greencircle}{\textcolor{Green}{\Huge{\ensuremath{\bullet}}}}

\newcommand{\greensquareS}{\raisebox{1.5pt}{\textcolor{Green}{\normalsize{\ensuremath{\blacksquare}}}}}
\newcommand{\greensquareSS}{\raisebox{1.5pt}{\textcolor{Green}{\footnotesize{\ensuremath{\blacksquare}}}}}
\newcommand{\bluecircleS}{\textcolor{blue}{\Large{\ensuremath{\bullet}}}}
\newcommand{\greencircleS}{\textcolor{Green}{\Large{\ensuremath{\bullet}}}}

\newcommand{\soc}{\ensuremath{\theta}\xspace}

@@

\usepackage{titlesec}

@@

\pagenumbering{gobble} 
% left justification:
\raggedright

%\usepackage{authblk}

% misc
\usepackage{graphicx}
\usepackage{babel}
\usepackage{amsmath}

\makeatletter
\makeatother

% Customisations specific to Philosophy of Science:

% paper size, margins, etc:
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1.5in,bmargin=1.5in,lmargin=1.25in,rmargin=1in}

@@

\renewcommand{\footnotesize}{\normalsize}

\usepackage{setspace} % load setspace before footmisc

\usepackage{footmisc}
\setlength{\footnotesep}{\baselineskip}

\usepackage{setspace,caption}
\AtBeginCaption{\doublespacing}

\renewcommand{\footnotelayout}{\doublespacing}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\normalsize}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

@@

%\titleformat{\subsection}
%       {\normalfont\fontfamily{phv}\fontsize{12}{17}\itshape}{\thesubsection}{1em}{}

\usepackage{titlesec}
\titleformat*{\section}{\bf}
\titleformat*{\subsection}{\it}
\titleformat*{\subsubsection}{\it}

% indentation
\setlength\parindent{12pt}

@@

\begin{abstract}
\normalsize
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals for decision making. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

\newpage

@@

In full generality, a choice mechanism associates decision situations with action choices. A crucial
part of a choice mechanism is the \emph{subjective representation} of the decision situation,
in particular the manner of forming preferences and beliefs about a possibly uncertain world (see
figure~\ref{fig:ChoiceMechanism} below). This paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness?
The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. A contribution of this paper is
to provide a case that questions of preference evolution should take
variability in uncertainty representation into account as well. As a point in case, we demonstrate
that if agents have imprecise probabilistic beliefs \citep[e.g.][]{gardsah82,levi74,walley96},
preference representations in terms of regret can be selected for, even though they may deviate
from the true, objective fitness that evolutionary selection operates on.

@@

\section{Rationality, Representations and Meta-games}
\label{sec:rati--subj}

@@

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. Adaptationism
aims to explain rational behavior by appeal to evolutionary considerations: DMs have acquired
choice mechanisms that have proved to be adaptive with respect to the variable environment
where they have evolved. A choice mechanism can be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to closely relate to the literature on evolution of preferences and to the philosophical debate
about the nature of rational choice in general, we here suggest to think of a choice mechanism
as a map from choice situations to action choices which includes an explicit level of
subjective representation of the situation, that we call \emph{subjective representation
  scheme} (see figure~\ref{fig:ChoiceMechanism}). Specifically, a subjective representation
scheme is a general way of forming preferences and beliefs about the choice situation, and we
are most interested in the question which subjective representations are better than others
from an evolutionary point of view.

@@

\section{Classic Choice Rules and Choice Mechanisms}
\label{sec:basic-notions}

@@

The case study under consideration strips the generality of a meta-game approach down to a perspicuous
comparison between four choice mechanisms, summarized in table~\ref{tab:CMs}.
% the cross-product of two different preference representations (veridical vs.~regret) and two
%different belief representations (precise vs.~imprecise). 
\begin{table}[t]
  \centering
  \begin{tabular}{cccc}
  \multicolumn{3}{c}{choice mechanism} \\  \cmidrule(r){1-3}
    \multicolumn{2}{c}{subjective representation} & action selection & corresponding classic rule \\  \cmidrule(r){1-2}
    preference type & belief type  \\ \midrule
    objective fitness & precise & maxmin EU &  Laplace rule \\
    objective fitness & imprecise & maxmin EU & maxmin \\ 
    regret & precise & maxmin EU & Laplace rule \\
    regret & imprecise & maxmin EU & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent)
    classic choice rules.}
  \label{tab:CMs}
\end{table}
\iffalse
This exclusive selection of choice mechanisms is motivated by a number of
considerations. Firstly, the set of conceivable choice mechanisms is vast. Most of that
conceivable vastness is utter nonsense, and some selection of plausible candidates is required in
any case. Secondly, a restriction to only a small number of suitable candidates will
help demonstrate more distinctly how a meta game approach may help structure philosophical
inquiry, which is our main goal. Thirdly, our case study adds to the recent literature on the
evolution of preferences by making evident the possibly non-trivial interaction between
preference and belief representations, all else equal. This is our secondary goal. 
Finally,
\fi 
The
four choice mechanisms,  
%that we look at here
although phrased in terms of preference and belief representations, map onto well-known classic
proposals for rational decision making (see table~\ref{tab:CMs}). In other words, there is also
motivation from historical relevance for our selection. To see this, we first recapitulate the
relevant classics ---to have a handy label, we call them \emph{classic rules}--- and then show
how they relate to our notion of choice mechanism.

\subsection{Classic Rules for Decision Making}
\label{sec:choice-rules}

@@

A strategic-form game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i: \bigtimes_{j \in N} A_j \rightarrow \mathbb{R}$ a payoff function for each player $i \in N$.
Part of this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(a_k, a_j) = \pi_2(a_j, a_k)$ for all $a_j,a_k \in A_1$. An example is given in
figure~\ref{coordgame1}. The numbers give the payoff for the player choosing a row act against
a co-player choosing a column act. Since the game is symmetric, it suffices to give just these
numbers.

Rational choice from maximization of expected utility, as given in
Definition~\ref{def:rationality}, applies to games as follows. To say which acts are rational,
it requires specifying a precise probabilistic belief of DM. Let $A_{-i} := \bigtimes_{j \neq i} A_j$ be
the set of all \emph{profiles} of acts for all players except $i$. If player $i$ has a belief
$\mu_i \in \Delta(A_{-i})$ about the behavior of his co-players, he chooses rationally only if
his choice $a^*_i$ maximizes expected utility
$\EU(a_i, \mu) := \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. For
the example in figure~\ref{coordgame1}, an expected utility maximizer would choose act $I$ only if he
believes that his co-player is going to choose act $I$ with a probability of at least
$\nicefrac{2}{3}$.

Perhaps the most well-known classic rule that does not require precise probabilistic
beliefs is the \emph{maxmin rule}. It prescribes to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios:
$\act^* \in \argmax_{a_i \in \Acts_i} \min_{a_{-i} \in \Acts_{-i}} \pi_i(a_{i},
a_{-i})$.
For the example in figure~\ref{coordgame1}, the minimum of both acts is 0: players choosing by maxmin
rule would be indifferent between acts $I$ and $II$.
